---
title: "🚧 WIP 🚧"
format: gfm
editor: visual
---

```{r}
#| label: "setup"
#| include: false

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

DataFusion Python API: https://arrow.apache.org/datafusion-python/

## What's this?

The main purpose of this repo is to see if [the savvy framework](https://github.com/yutannihilation/savvy) has enough features for mildly complex usages. Especially,

-   to check if savvy can convert an external pointer or ALTREP that an other package defines. e.g. <https://arrow.apache.org/datafusion-python/>
-   to figure out what async support is supposed to be e.g. \<https://github.com/yutannihilation/savvy-async-poc?

## Usages

```{r}
library(datafusionr)

x <- arrow::as_record_batch(head(mtcars))

ctx <- datafusion_session_context()
d <- ctx$create_data_frame(as_array_stream(x), table_name = "mtcars")
d
```

### `limit()`

```{r}
d |> 
  datafusion_limit(2)
```


### `select_columns()`

```{r}
d |>
  datafusion_select_columns("cyl", "mpg")
```
### `select()`

```{r}
d |> 
  datafusion_select(
    (col("mpg") / col("cyl"))$alias("foo")
  )
```

