---
title: "🚧 WIP 🚧"
format: gfm
editor: visual
---

```{r}
#| label: "setup"
#| include: false

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<!-- badges: start -->
[![R-CMD-check](https://github.com/yutannihilation/datafusionr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/yutannihilation/datafusionr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

## What's this?

The main purpose of this repo is to see if [the savvy framework](https://github.com/yutannihilation/savvy) has enough features for mildly complex usages. Especially,

-   to check if savvy can convert an external pointer or ALTREP that an other package defines. e.g. <https://arrow.apache.org/datafusion-python/>
-   to figure out what async support is supposed to be e.g. \<https://github.com/yutannihilation/savvy-async-poc?

## Usages

### DataFrame

```{r}
library(datafusionr)

x <- arrow::as_record_batch(head(mtcars))

ctx <- datafusion_session_context()
d <- ctx$create_data_frame(as_array_stream(x), table_name = "mtcars")
d
```

`datafusion_collect()` can convert the object into a bare `data.frame`.

```{r}
d |> 
  datafusion_collect()
```


#### `limit()`

```{r}
d |> 
  datafusion_limit(2)
```


#### `select_columns()`

```{r}
d |>
  datafusion_select_columns("cyl", "mpg")
```
#### `select()`

```{r}
d |> 
  datafusion_select(
    col("mpg")$alias("MPG"),  # rename
    sqrt(col("mpg") / lit(2)) # arbitrary expression
  )
```

### Parquet file

```{r}
ctx_parquet <- datafusion_session_context()

parquet_file <- system.file("extdata/mtcars.parquet", package = "datafusionr")
ctx_parquet$register_parquet("mtcars", parquet_file)

d <- ctx_parquet$sql("select cyl, mpg from mtcars limit 5")
d
```

### CSV file

```{r}
ctx_csv <- datafusion_session_context()

csv_file <- system.file("extdata/mtcars.csv", package = "datafusionr")
ctx_csv$register_csv("mtcars", csv_file)

d <- ctx_csv$sql("select cyl, mpg from mtcars limit 5")
d
```
